---
title: "Energy Production and Consumption Analysis"
output: html_notebook
---

# Energy Production and Consumption Analysis (R Version)

This notebook replicates the energy production and consumption analysis performed in the original Python Jupyter Notebook. It includes data preprocessing, visualization, Multi-Criteria Decision Making (MCDM) analysis, and Machine Learning models for classification and regression.

## 1. Importing Dependencies

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(caret)
library(randomForest)
library(e1071)
library(class)
library(forecast)
library(xgboost)
library(corrplot)
library(scales)

# For MCDM (Manual Implementation or using a library)
# library(MCDM) 
```

## 2. Loading and Preprocessing Datasets

```{r load_data}
# Load the datasets
hydro <- read_csv("data/US_Geo_Hydro.csv")
solar <- read_csv("data/US_Solar.csv")
wind <- read_csv("data/US_Wind.csv")

# Rename columns and add energy type
rename_cols <- function(df, type) {
  df %>%
    rename(area = area_developable_sq_km,
           capacity = capacity_ac_mw) %>%
    mutate(energy_type = type)
}

hydro <- rename_cols(hydro, "Hydro")
solar <- rename_cols(solar, "Solar")
wind <- rename_cols(wind, "Wind")

# Combine datasets and handle missing values
combined_df <- bind_rows(hydro, solar, wind) %>%
  drop_na(latitude, longitude, area, capacity)

head(combined_df)
```

## 3. Visualization

### Number of Sites by Energy Type
```{r vis_count}
ggplot(combined_df, aes(x = energy_type, fill = energy_type)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Number of Sites by Energy Type", x = "Energy Type", y = "Count") +
  theme_minimal()
```

### Capacity Distribution
```{r vis_dist_capacity}
ggplot(combined_df, aes(x = capacity, fill = energy_type)) +
  geom_histogram(bins = 50, alpha = 0.6, position = "identity") +
  scale_x_log10() +
  labs(title = "Capacity Distribution by Energy Type", x = "Capacity (AC MW) - Log Scale", y = "Frequency") +
  theme_minimal()
```

### Site Locations
```{r vis_map}
ggplot(combined_df, aes(x = longitude, y = latitude, color = energy_type)) +
  geom_point(alpha = 0.4, size = 0.5) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Site Locations by Energy Type", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

## 4. Multi-Criteria Decision Making (MCDM)

Implementation of TOPSIS ranking based on capacity and area.

```{r topsis_ranking}
# Simple TOPSIS implementation
topsis <- function(matrix, weights, criteria) {
  # Normalize matrix
  norm_matrix <- sweep(matrix^2, 2, colSums(matrix^2), "/") %>% sqrt()
  # Weighted matrix
  weighted_matrix <- sweep(norm_matrix, 2, weights, "*")
  
  # Ideal best and worst
  ideal_best <- apply(weighted_matrix, 2, max)
  ideal_worst <- apply(weighted_matrix, 2, min)
  
  # Distances
  dist_best <- sqrt(rowSums(sweep(weighted_matrix, 2, ideal_best, "-")^2))
  dist_worst <- sqrt(rowSums(sweep(weighted_matrix, 2, ideal_worst, "-")^2))
  
  # Score
  score <- dist_worst / (dist_best + dist_worst)
  return(score)
}

# Apply TOPSIS
mcdm_data <- combined_df %>% select(capacity, area) %>% as.matrix()
weights <- c(0.6, 0.4)
scores <- topsis(mcdm_data, weights, c(1, 1))
combined_df$topsis_score <- scores

combined_df %>% 
  arrange(desc(topsis_score)) %>% 
  select(latitude, longitude, capacity, area, topsis_score) %>% 
  head(5)
```

## 5. Classification (Energy Type)

```{r classification}
# Select features
features <- combined_df %>% 
  select(latitude, longitude, area, capacity) %>% 
  mutate(energy_type = as.factor(combined_df$energy_type))

# Split data
set.seed(42)
trainIndex <- createDataPartition(features$energy_type, p = .8, list = FALSE)
train_data <- features[trainIndex,]
test_data  <- features[-trainIndex,]

# Random Forest Model
model_rf <- randomForest(energy_type ~ ., data = train_data, ntree = 100)
preds_rf <- predict(model_rf, test_data)

# Confusion Matrix
confusionMatrix(preds_rf, test_data$energy_type)
```

## 6. Regression (Capacity Prediction)

```{r regression}
# Linear Regression
model_lm <- lm(capacity ~ area + latitude + longitude, data = train_data)
summary(model_lm)

# Predictions
preds_lm <- predict(model_lm, test_data)
RMSE(preds_lm, test_data$capacity)
```

## 7. Time Series Forecasting (Consumption)

```{r time_series}
# Load consumption data
cons_df <- read_excel("data/US_Prod_and_Cons.xlsx")
# Ensure Month is Date
cons_df$Month <- as.Date(cons_df$Month)

# Create Time Series object
ts_data <- ts(cons_df$`Total Renewable Energy Consumption`, frequency = 12, start = c(year(min(cons_df$Month)), month(min(cons_df$Month))))

# ARIMA Model
fit_arima <- auto.arima(ts_data)
forecast_arima <- forecast(fit_arima, h = 12)

# Plot Forecast
plot(forecast_arima, main = "Renewable Energy Consumption Forecast")
```

## Conclusion

This R notebook successfully replicates the core analysis steps from the Python implementation, including data processing, visualization, site ranking, and predictive modeling.
